---
title: "Data608 - Homework 1"
author: "Alec"
date: "9/11/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results='hide', message=FALSE}
library(tidyverse)
library(ggplot2)
library(psych)
```


**Principles of Data Visualization and Introduction to ggplot2**

I have provided you with data about the 5,000 fastest growing companies in the US, as compiled by Inc. magazine. lets read this in:

```{r}
inc <- read.csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module1/Data/inc5000_data.csv", header= TRUE)
```

And lets preview this data:

```{r}
head(inc)
```

```{r}
summary(inc)
```

Think a bit on what these summaries mean. Use the space below to add some more relevant non-visual exploratory information you think helps you understand this data:

```{r}
# glimpse is helpful if there are lots of features. However, in this case it does not provide much utility
glimpse(inc)
```

```{r}
# describe() is a powerful summary tool, giving the user everything in summary() and more
describe(inc)
```

```{r}
# the last thing I typically check is for null values
colSums(is.na(inc))
```

```{r}
# Lets check out the rows with missing entries
inc %>%
  filter(is.na(Employees))
```



One can approach the case of missing values in a few ways. Imputation is one option, and removing the rows is another. In this case, because the revenues of the companies in question vary widely (they are not all small companies) it could be appropriate to impute with the mean.

```{r}
inc$Employees[is.na(inc$Employees)] <- mean(inc$Employees,na.rm = TRUE)
```


## Question 1

Create a graph that shows the distribution of companies in the dataset by State (ie how many are in each state). There are a lot of States, so consider which axis you should use. This visualization is ultimately going to be consumed on a 'portrait' oriented screen (ie taller than wide), which should further guide your layout choices.

```{r}
aggregated <- inc %>%
  group_by(State) %>%
  count(State) %>%
  arrange(desc(n))

aggregated %>%
  ggplot(aes(y=reorder(State, n), x=n)) +
  geom_bar(stat="identity") +
  xlab("Count") + 
  ylab("State")

```



## Quesiton 2

Lets dig in on the state with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries. Create a plot that shows the average and/or median employment by industry for companies in this state (only use cases with full data, use R's `complete.cases()` function.) In addition to this, your graph should show how variable the ranges are, and you should deal with outliers.

```{r}
# New York is the third largest by # companies
ny_state <- inc %>% filter(State=="NY")
summary(ny_state)
```

```{r}
# plotting NY state
ny_state %>% 
  filter(complete.cases(.)) %>% # complete cases only
  group_by(Industry) %>% 
  select(Industry, Employees) %>%
  ggplot(mapping=aes(x=Industry, y=Employees)) +
  geom_boxplot()

```

This will not do. There are many high outliers which are skewing the image. Additionally, the text at the bottom is illegible.

```{r}
# plotting NY state
ny_state %>% 
  filter(complete.cases(.)) %>% # complete cases only
  group_by(Industry) %>% 
  select(Industry, Employees) %>%
  ggplot(mapping=aes(x=Industry, y=Employees)) +
  geom_boxplot(outlier.shape=NA) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  coord_cartesian(ylim = c(0, 1500))

```



## Question 3

Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart that makes this information clear. Once again, the distribution per industry should be shown.

```{r}
# first create the new variable
inc$rev_per_employee = inc$Revenue / inc$Employees
```


```{r}
head(inc, 20)
```


```{r}
# plot with facet wrap
ggplot(inc, aes(rev_per_employee)) +
  geom_density() +
  facet_wrap(~Industry)
```

```{r}
ny_state %>%
  group_by(Industry) %>%
  summarize(total_rev = sum(Revenue), 
            total_emp = sum(Employees), 
            rev_per_emp = total_rev/total_emp) %>%
  arrange(desc(rev_per_emp)) %>%
  na.omit() %>%
  ggplot(aes(x=reorder(Industry, rev_per_emp), y=rev_per_emp)) +
  geom_bar(stat="identity") +
  coord_flip()
```








