---
title: "608 Homework 3 - Shiny"
author: "Alec"
date: "10/9/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)  
library(dplyr)    
library(tidyr)    
library(tibble)      
library(reshape2)   
library(stringr) 

library(plotly)
library(shiny)
library(rsconnect)
```


```{r}
# authenticate

rsconnect::setAccountInfo(name='man-of-moose',
			  token='38AA0D6DCC7CB17D6504F090A7FE84DD',
			  secret='fA+ocz1KAwE8eC3RTevfDGRCHUBq8GTu+tbHR1x3')
```



```{r}
data <- read_csv("https://raw.githubusercontent.com/charleyferrari/CUNY_DATA_608/master/module3/data/cleaned-cdc-mortality-1999-2010-2.csv")
```

```{r}
data
```

## Section 1

As a researcher, you frequently compare mortality rates from particular causes across different States. You need a visualization that will let you see (for 2010 only) the crude mortality rate, across all States, from one cause (for example, Neoplasms, which are effectively cancers). Create a visualization that allows you to rank States by crude mortality for each cause of death.

```{r}
cod = "Neoplasms"
pp1 <- data %>%
  dplyr::filter(., Year == "2010" & ICD.Chapter == cod) %>% # Filter by year and COD
  arrange(desc(State), Crude.Rate)                   # Reverse the State order to flip axes

head(pp1)
```

```{r}
chart1 <- pp1 %>%
  plot_ly(x = ~pp1$Crude.Rate, y = pp1$State, type="bar", orientation="h") %>%
  layout(
          title= list(text=paste0(cod,"\ncause-based 2010 State Crude Mortality Rates"),
                      font=list(size = 10)),
    xaxis=list(title="Crude Rate"),
    yaxis=list(title="States",
      categoryorder = "array",
      categoryarray = rev(~State))                    # Flip axes
  )

subplot(chart1) 
```

```{r}
# Shiny

sidebarPanel(
  selectInput("cod", label = "Cause of Death:",
              choices = data$ICD.Chapter)
)

mainPanel( 
  plotlyOutput('chart2') 
)

renderPlot({
  pp1 <- data %>%
    filter(., Year == "2010" & ICD.Chapter == input$cod) %>% # Filter by year and COD
    arrange(desc(State), Crude.Rate)                   # Reverse the State order to flip axes
  
  chart1 <- pp1 %>%
    plot_ly(x = ~pp1$Crude.Rate, y = pp1$State, type="bar", orientation="h") %>%
    layout(
          title= list(text=paste0(input$cod,"\ncause-based 2010 State Crude Mortality Rates"),
                      font=list(size = 10)),
          xaxis=list(title="Crude Rate"),
          yaxis=list(title="States",
            categoryorder = "array",
            categoryarray = rev(~State))               # Flip axes
        )
  
  subplot(chart1)    
})
```

```{r}
#rsconnect::deployApp('/Users/alecmccabe/Desktop/Masters Program/DATA 608/DATA608/homework3/homework3_rmd.Rmd')
```


